---
title: "Assessment task part 1"
author: "Akanksha and Asad"
date: '2022-05-20'
output: html_document
---

##PART 1
## Source: The code for this Markdown script can be found on our Repository on Github, The link is provided below:
## Source = https://github.com/Akku-05/Assessment-Task

##Question 1

##Dowload the gene expression data and change the row names to gene expression name
```{r, part 1}
#Downloading the gene expression data provided on the unit site
#gene expression data

URL= "https://raw.githubusercontent.com/markziemann/SLE712_files/master/assessment_task3/bioinfo_asst3_part1_files/gene_expression.tsv"
gene= "gene_expression.tsv"
download.file(URL,destfile=gene)
#Reading the .tsv data into R
GE <- read.delim("gene_expression.tsv")

#Changing row names to gene expression name.
colnames(GE)
df <- data.frame(GE)
rownames(df) <- df$Name_Description
df$Name_Description=NULL

# subsetting first 6 genes
df[1:6,]

#formatting table
library(kableExtra)
df[1:6,]

head(df) %>%
  kbl(caption = "gene expression data table") %>%
  kable_paper(full_width = F, html_font = "Arial")

```

##QUESTION 2

##Make a new column for mean of gene expression and find the values for top 6 genes
```{r,question2}

colMeans(df[1:6,1:3]) 

rowMeans(df[1:6,1:3])

df$mean <- rowMeans(df)
A <- df[1:6,]
GeneMean <- rowMeans(A)
head(GeneMean) %>%
  kbl(caption = "gene expression mean table") %>%
  kable_paper(full_width = F, html_font = "Arial")
```

##QUESTION3

##10 gene with highest gene expression
```{r, QUESTION3}

# Defining variable B as mean in decreasing order
B <-order(df$mean, decreasing=TRUE)
HighestG <-B[1:10] 
head(HighestG) # Highest mean expression
```

##QUESTION 4

##Number of genes with mean less that 10
```{r, question 4}
df$mean < 10
Mean_filtered <- df[which(df$mean < 10),]
nrow(Mean_filtered)
```

##Question 5

##Histogram of mean values
```{r, question 5}
hist(df$mean, xlab = "mean", ylab = "expression", main = "Gene Expression")
```

##Question 6

##Downloading growth data csv file into R and determining column names
```{r}
URL= "https://raw.githubusercontent.com/markziemann/SLE712_files/master/assessment_task3/bioinfo_asst3_part1_files/growth_data.csv"
growth= "growth_data.csv"
download.file(URL,destfile=growth)
GD <-read.csv("growth_data.csv")
colnames(GD)
```

##Question 7

##Mean and standard deviation to tree circumferences at the start and end of the study for both sites
```{r, question 7}
NE<-subset(GD,Site =="northeast")
SW <- subset(GD,Site =="southwest")
NE2005 <- NE$Circumf_2005_cm
NE2020 <- NE$Circumf_2020_cm
SW2005 <- SW$Circumf_2005_cm
SW2020 <- SW$Circumf_2020_cm
mean(NE2005)
mean(NE2020)
mean(SW2005)
mean(SW2020)
a <- sd(NE2005)
b <-sd(NE2020)
c <-sd(SW2005)
d <-sd(SW2020)
```

##Question 8

##Boxplot for tree circumferences of both the sites from start and end of study
```{r, Question 8}
boxplot(NE2005,NE2020,SW2005,SW2020, main= "Growth of trees", at=c(1,2,3,4),
        names= c("NE2005","NE2020","SW2005","
 
                                 SW2020"), ylab="Growth(cm)")
grid()
```

##Question 9

##Mean Growth over the 10 years at each site
```{r, question 9}
NE2010 <- NE$Circumf_2010_cm

SW2010 <- SW$Circumf_2010_cm
growthrateNE <- ((NE2020-NE2010)/NE2010) #growth rate = ((present-past)growth/past growth)
growthrateSW <- ((SW2020-SW2010)/SW2010)
SWMean10 <- mean(growthrateSW)
NEMean10 <-mean(growthrateNE)
NEMean10      #Mean for northeast region for past 10 years
SWMean10     #Mean for southwest region for past 10 years
MeanGrowth <- mean(growthrateNE+growthrateSW)
MeanGrowth      #Mean growth rate of both the site for past 10 years
```

##Question 10

##T.test and wilcox.test to estimate P. value for difference in growth at 3 sites
```{r, question 10}
res <- t.test (growthrateNE,growthrateSW)  #Defining res as t.test for growth rate at both the sites
p.value <-res$p.value # p value for growth rate
HEADER <- paste("P Value:", signif(p.value,3))

p.value2 <-wilcox.test(growthrateNE,growthrateSW) # p value for growth rate
HEADER <- paste("P Value:", signif(p.value,3))

HEADER

```
