---
title: "Part 2"
author: "Akanksha and Asad"
date: '2022-06-01'
output: html_document
---

##Part 2 for Bioinformatics Report

## Source: The code for this Markdown script can be found on our Repository on Github, The link is provided below:
## Source = https://github.com/Akku-05/Assessment-Task

##Question 1
Downloading coding sequeneces for E.coli and S.bacterium. 
```{r, question 1}

library("seqinr")
library("kableExtra")
library("R.utils")

#QUESTION 1
#downloading E.coli coding DNA sequence

URL="http://ftp.ensemblgenomes.org/pub/bacteria/release-53/fasta/bacteria_0_collection/escherichia_coli_str_k_12_substr_mg1655_gca_000005845/cds/Escherichia_coli_str_k_12_substr_mg1655_gca_000005845.ASM584v2.cds.all.fa.gz"
download.file(URL,destfile = "ecoli_cds.fa.gz")
gunzip("ecoli_cds.fa.gz")


#downloading Saprospirales bacterium (GCA_003448025) coding DNA sequence

URL1="http://ftp.ensemblgenomes.org/pub/bacteria/release-53/fasta/bacteria_58_collection/saprospirales_bacterium_gca_003448025/cds/Saprospirales_bacterium_gca_003448025.ASM344802v1.cds.all.fa.gz"
download.file(URL1,destfile="Sbac_cds.fa.gz")
gunzip("Sbac_cds.fa.gz")

# Calculating the no. of coding sequences present

#listing the length of cds for E.coli and S.bacterium
cds <- seqinr::read.fasta("ecoli_cds.fa")
str(head(cds))
cds2 <- seqinr:: read.fasta("Sbac_cds.fa")

CS_E.coli <- length(cds)   #coding seq for E.coli calculating the no. of coding sequences
CS_S <- length(cds2) # coding seq for S.bacterium calculating the no. of coding sequences

# Calculating the Coding DNA present in both the organisms
len1 <- as.numeric(summary(cds)[,1])
len2 <- as.numeric(summary(cds2)[,1])

sum(len1)
sum(len2)


length(unlist(cds[1:3]))
length(unlist(cds2[1:3]))



dna1 <- unlist(cds)
dna2 <- unlist(cds2)
head(dna1)
head(dna2)



```
The number of coding sequences is comparitively more for S.bacterium as compared to E.coli.



##Question 2

##Length of all coding sequences
```{r, question 2}
# Mean and Median for E.coli
head(summary(cds))
str(summary(cds))

head( summary(cds)[,1])
len3 <- sapply(X=cds,FUN=length) #calculating the
median(len3)
mean(len3)
#Mean and median for S.bacterium
head(summary(cds2))
str(summary(cds2))
head( summary(cds2)[,1])
len4 <- sapply(X=cds2,FUN=length)
median(len4)
mean(len4)

#Boxplot of coding sequences
boxplot(len3,len4,main= "Length of Coding Sequences", at=c(1,2),
        names= c("E.Coli","S.bacterium"), ylab="Length")

```
##Describe differences between both the organisms 
The length of the coding sequence of E.coli is less than that of S.bacterium. Coding sequence length is importantly associated with biological timing. The smaller sequences produce smaller proteins faster, and these proteins often play a part in the regulation of longer proteins, which are expressed much later into the response.Meaning, that the small proteins in E.coli will be produced faster than S.bacterium.
This difference in coding sequence length can be shown by mean,median and boxplot generated.


##Question 3

```{r, question 3}
#calculating dna composition frquency for E.coli
GC(cds[[1]])
count(cds[[1]],1)
count(cds[[1]],2)
count(cds[[1]],3)
summary(cds[1:3])
sum(sapply(cds[1:3],length))
length(unlist(cds[1:3]))
dna1 <- unlist(cds)
GC(dna1)
dna_composition <- count(dna1,1)
#Calculating dna composition frequency for S.bacterium

GC(cds2[[1]])
count(cds2[[1]],1)
count(cds2[[1]],2)
count(cds2[[1]],3)
summary(cds2[1:3])
sum(sapply(cds2[1:3],length))
length(unlist(cds2[1:3]))
dna2 <- unlist(cds2)
GC(dna2)
dna_composition2 <- count(dna2,1)

# Bar plot fpr DNA composition
barplot(dna_composition,xlab="nucleotides",ylab="frequency", main="E.Coli DNA composition")

barplot(dna_composition2,xlab="nucleotides",ylab="frequency", main="S.bacterium DNA composition")



#Calculating protein sequence for E.coli
translate(cds[[1]])
prot <- lapply(cds,translate)
prot[[1]]
aa <-unique(prot[[2]])
aa <- aa[aa !="*"]

protein <- unlist(prot)

protein_composition <- count(protein,alphabet=aa,wordsize=1)
protein_composition

protseq<- count(protein,alphabet=aa,wordsize=1)

#Calculating protein sequence for S.bacterium
translate(cds2[[1]])
prot2 <- lapply(cds2,translate)
prot2[[1]]
aa2 <-unique(prot[[2]])
aa2 <- aa2[aa2 !="*"]

protein2 <- unlist(prot2)

protein_composition2 <- count(protein2,alphabet=aa2,wordsize=1)


protseq2<- count(protein2,alphabet=aa2,wordsize=1)

#Barplots for Protein Composition

barplot(protseq,xlab="Protein composition",ylab="frequency", main="E.Coli Protein composition")
barplot(protseq2,xlab="Protein composition",ylab="frequency", main="S.bacterium Protein composition")

```
##Describe any differences between two organisms.
The frquency of DNA bases in E.coli is fairly even however there are more guanine more frequently expressed than the three bases. The frequency of DNA bases in S.bacterium is more staggered, with adenocine being the most frequent and cytosine the being least frequent.
The frquency of protein seuence in E.coli can be visualised by the barplot, which shows leucine, alanine and glycine as the most frequent amino acid and cystine, tryptophan and histadine being the least frequent amino acid.
The frequency of DNA bases in E.coli is fairly even however there are more guanine more frequently expressed than the three bases. The frequency of DNA bases in S.bacterium is more staggered, with adenosine being the most frequent and cytosine the being least frequent thereby showing less differnece in the protein sequence.

##Question 4

```{r, question 4}
uco(cds[[2]])
codon_usage <- uco(cds[[2]],index = "rscu",as.data.frame =TRUE)

codon_usage
codon_usage%>%
  kbl()%>%
  kable_paper(full_width = F, html_font = "Arial")

plot(codon_usage, main="Codon bias for E.coli")

#Condon usage for S.bacterium
uco(cds2[[2]])
codon_usage2 <- uco(cds2[[2]],index = "rscu",as.data.frame =TRUE)

codon_usage2
codon_usage2%>%
  kbl()%>%
  kable_paper(full_width = F, html_font = "Arial")
plot(codon_usage2, main="Codon Bias for S.bacterium")

```
#Describe any differneces with repect to codon usage bias\
Amino acid such as almine, valine and leucine are considered hydrophobic amino acid. When comparing E.coli protein sequence with S.bacterium protein sequence, the elanin concentration is higher in E.coli as well as the valine concentartion is also higher in E.coli than the S.bacterium, which makes the E.coli isb more than the S.bacterium.
Amino acid such as serine, threonine and gultamine are considered as polar amino acid. In E.coli protein, serine has more concentration than in S.bacterium, which makes E.coli more polar protein sequence.





##Question 5

```{r, question 5}

#Kmer profiling for E.coli
protein <- unlist(prot)

Kmer1 <- count(protein,wordsize=3,alphabet=aa)
Kmer2 <- count(protein,wordsize=4,alphabet=aa)
Kmer3 <- count(protein,wordsize=5,alphabet=aa)



head(Kmer1)
aa[aa !="*"]

myfreq1 <- count(protein,wordsize=3,alphabet=aa,freq=TRUE)
myfreq2 <- count(protein,wordsize=4,alphabet=aa,freq=TRUE)
myfreq3 <- count(protein,wordsize=5,alphabet=aa,freq=TRUE)


#10 overexpressed kmers of length 3 for E.coli
head(myfreq1,n=10) #Kmer profiling for Kmers length 3

#10 underexpressed kmers of length 3 for E.coli
tail(myfreq1,n=10)

#10 overexpressed kmers of length 4 for E.coli
head(myfreq2,n=10) #Kmer profiling for Kmers length 4

#10 underexpressed kmers of length 5 for E.coli
tail(myfreq2,n=10)

#10 overexpressed kmers of length 3 for E.coli
head(myfreq3,n=10) #Kmer profiling for Kmers length 5

#10 underexpressed kmers of length 5 for E.coli
tail(myfreq3,n=10)

#Kmer profiling for S.bacterium
protein2 <- unlist(prot2)

Kmer4<- count(protein2,wordsize =3,alphabet=aa2)
Kmer5<- count(protein2,wordsize = 4,alphabet=aa2)
Kmer6<- count(protein2,wordsize = 5,alphabet=aa2)


head(Kmer4)
aa[aa !="*"]

myfreq4<- count(protein2,wordsize =3,alphabet=aa2,freq = TRUE)
myfreq5 <- count(protein2,wordsize = 4,alphabet=aa2,freq=TRUE)
myfreq6 <- count(protein2,wordsize = 5,alphabet=aa2,freq=TRUE)

#10 overexpressed kmers of length 3 for S.bacterium
head(myfreq4,n=10) #Kmer profiling for Kmers length 3

#10 underexpressed kmers of length 3 for S.bacterium
tail(myfreq4,n=10)

#10 overexpressed kmers of length 4 for S.bacterium
head(myfreq5,n=10) #Kmer profiling for Kmers length 4

#10 underexpressed kmers of length 4 for S.bacterium
tail(myfreq5,n=10)

#10 overexpressed kmers of length 5 for S.bacterium
head(myfreq6,n=10) #Kmer profiling for Kmers length 5

#10 underexpressed kmers of length 4 for S.bacterium
tail(myfreq6,n=10)

plot(myfreq1,main="Kmer Profiling for E.Coli",xlab = "Kmers of Length 3",ylab="Frequency")

plot(myfreq4,main="Kmer Profiling for S.bacterium",xlab = "Kmers of Length 3",ylab="Frequency")

```


